---
title: "read_depth.rmd"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/Jacob/Desktop/Menzies/Knowlesi/Pipeline/Pk_Pipeline/')
```

# Packages 

```{r}
library(tidyverse)
library(janitor)
```
# Direct Alignments 

# Samples SS13 and ZB31

SS average coverage = 2.6, 3.4, 28.3, 3.4, 3.3, 7.9, 274.2, 62.2, 3.1, 61.3, 9.0, 0.035, 5.5, 19.2
SS bases with coverage (%) = 54.1, 54.4, 70.1, 47.0, 56.2, 80.7, 99.8, 89.0, 46.6, 53.5, 75.3, 1.9, 57.3, 77.4

ZB average coverage = 124.0, 115.6, 123.6, 137.6, 105.9, 139.7, 42, 106.2, 129.1, 97.4, 140.4, 4.5, 176.8, 33.9
ZB bases with coverage (%) = 84.7, 75.5, 83.4, 80.9, 75.7, 91.4, 99.5, 94.3, 79.1, 83.1, 85.1, 14.6, 86.6, 79.6

```{r}
data.frame(SS_average_coverage_depth = c(2.6, 3.4, 28.3, 3.4, 3.3, 7.9, 274.2, 62.2, 3.1, 61.3, 9.0, 0.035, 5.5, 19.2),
                         SS_bases_with_coverage = c(54.1, 54.4, 70.1, 47.0, 56.2, 80.7, 99.8, 89.0, 46.6, 53.5, 75.3, 1.9, 57.3, 77.4),
                         ZB__average_coverage_depth = c(124.0, 115.6, 123.6, 137.6, 105.9, 139.7, 42, 106.2, 129.1, 97.4, 140.4, 4.5, 176.8, 33.9),
                         ZB_bases_with_coverage = c(84.7, 75.5, 83.4, 80.9, 75.7, 91.4, 99.5, 94.3, 79.1, 83.1, 85.1, 14.6, 86.6, 79.6)) %>% 
  add_column(contig = 1:14) %>% 
  mutate(contig = as.factor(contig)) %>% 
  summarise_all(mean)
```

```{r}
coverage_plot1 <- data.frame(average_coverage_depth = c(2.6, 3.4, 28.3, 3.4, 3.3, 7.9, 274.2, 62.2, 3.1, 61.3, 9.0, 0.035, 5.5, 19.2),
           bases_with_coverage = c(54.1, 54.4, 70.1, 47.0, 56.2, 80.7, 99.8, 89.0, 46.6, 53.5, 75.3, 1.9, 57.3, 77.4)) %>% 
           add_column(data = "SS") %>% 
           add_column(contig = 1:14) %>% 
           mutate(contig = as.factor(contig)) %>% 
  rbind(
    data.frame(average_coverage_depth = c(124.0, 115.6, 123.6, 137.6, 105.9, 139.7, 42, 106.2, 129.1, 97.4, 140.4, 4.5, 176.8, 33.9),
           bases_with_coverage = c(84.7, 75.5, 83.4, 80.9, 75.7, 91.4, 99.5, 94.3, 79.1, 83.1, 85.1, 14.6, 86.6, 79.6)) %>% 
           add_column(data = "ZB") %>% 
           add_column(contig = 1:14) %>% 
           mutate(contig = as.factor(contig))) %>% 
           mutate_if(is.character, as.factor) %>% 
  ggplot(mapping = aes(x = contig, y = average_coverage_depth, colour = data)) +
  geom_col() + 
  facet_wrap(~data)
```

```{r}
bases_plot1 <- data.frame(average_coverage_depth = c(2.6, 3.4, 28.3, 3.4, 3.3, 7.9, 274.2, 62.2, 3.1, 61.3, 9.0, 0.035, 5.5, 19.2),
           bases_with_coverage = c(54.1, 54.4, 70.1, 47.0, 56.2, 80.7, 99.8, 89.0, 46.6, 53.5, 75.3, 1.9, 57.3, 77.4)) %>% 
           add_column(data = "SS") %>% 
           add_column(contig = 1:14) %>% 
           mutate(contig = as.factor(contig)) %>% 
  rbind(
    data.frame(average_coverage_depth = c(124.0, 115.6, 123.6, 137.6, 105.9, 139.7, 42, 106.2, 129.1, 97.4, 140.4, 4.5, 176.8, 33.9),
           bases_with_coverage = c(84.7, 75.5, 83.4, 80.9, 75.7, 91.4, 99.5, 94.3, 79.1, 83.1, 85.1, 14.6, 86.6, 79.6)) %>% 
           add_column(data = "ZB") %>% 
           add_column(contig = 1:14) %>% 
           mutate(contig = as.factor(contig))) %>% 
           mutate_if(is.character, as.factor) %>% 
  ggplot(mapping = aes(x = contig, y = bases_with_coverage, colour = data)) +
  geom_col() + 
  facet_wrap(~data)
```


# Samples SS10 and ZB09


SS bases with coverage (%) = 38.6, 33.0, 52.2, 32.0, 37.6, 70.0, 99.1, 77.4, 27.0, 38.3, 58.3, 0.8, 46.3, 63.3
SS average coverage = 1.2,1.5, 6.1, 1.6, 1.2, 9.4, 159.3, 13.5, 1.1, 9.5, 2.6, 0.01, 4.6, 8.6

ZB bases with coverage (%) = 66.6, 55.8, 62.7, 65.0, 56.6, 77.0, 70.5, 76.5, 60.0, 64.0, 73.8, 9.4, 62.8, 50.6
ZB average coverage = 9.9, 6.6, 11.5, 11.0, 7.7, 13.4, 4.3, 7.8, 8.8, 9.0, 13.4, 0.3, 17.6, 3.0

```{r}
data.frame(SS_average_coverage_depth = c(1.2,1.5, 6.1, 1.6, 1.2, 9.4, 159.3, 13.5, 1.1, 9.5, 2.6, 0.01, 4.6, 8.6),
                         SS_bases_with_coverage = c(38.6, 33.0, 52.2, 32.0, 37.6, 70.0, 99.1, 77.4, 27.0, 38.3, 58.3, 0.8, 46.3, 63.3),
                         ZB__average_coverage_depth = c(9.9, 6.6, 11.5, 11.0, 7.7, 13.4, 4.3, 7.8, 8.8, 9.0, 13.4, 0.3, 17.6, 3.0),
                         ZB_bases_with_coverage = c(66.6, 55.8, 62.7, 65.0, 56.6, 77.0, 70.5, 76.5, 60.0, 64.0, 73.8, 9.4, 62.8, 50.6)) %>% 
  add_column(contig = 1:14) %>% 
  mutate(contig = as.factor(contig)) %>% 
  summarise_all(mean)
```

```{r}
coverage_plot2 <- data.frame(average_coverage_depth = c(1.2,1.5, 6.1, 1.6, 1.2, 9.4, 159.3, 13.5, 1.1, 9.5, 2.6, 0.01, 4.6, 8.6),
           bases_with_coverage = c(38.6, 33.0, 52.2, 32.0, 37.6, 70.0, 99.1, 77.4, 27.0, 38.3, 58.3, 0.8, 46.3, 63.3)) %>% 
           add_column(data = "SS") %>% 
           add_column(contig = 1:14) %>% 
           mutate(contig = as.factor(contig)) %>% 
  rbind(
    data.frame(average_coverage_depth = c(9.9, 6.6, 11.5, 11.0, 7.7, 13.4, 4.3, 7.8, 8.8, 9.0, 13.4, 0.3, 17.6, 3.0),
           bases_with_coverage = c(66.6, 55.8, 62.7, 65.0, 56.6, 77.0, 70.5, 76.5, 60.0, 64.0, 73.8, 9.4, 62.8, 50.6)) %>% 
           add_column(data = "ZB") %>% 
           add_column(contig = 1:14) %>% 
           mutate(contig = as.factor(contig))) %>% 
           mutate_if(is.character, as.factor) %>% 
  ggplot(mapping = aes(x = contig, y = average_coverage_depth, colour = data)) +
  geom_col() + 
  facet_wrap(~data)
```

```{r}
bases_plot2 <- data.frame(average_coverage_depth = c(1.2,1.5, 6.1, 1.6, 1.2, 9.4, 159.3, 13.5, 1.1, 9.5, 2.6, 0.01, 4.6, 8.6),
           bases_with_coverage = c(38.6, 33.0, 52.2, 32.0, 37.6, 70.0, 99.1, 77.4, 27.0, 38.3, 58.3, 0.8, 46.3, 63.3)) %>% 
           add_column(data = "SS") %>% 
           add_column(contig = 1:14) %>% 
           mutate(contig = as.factor(contig)) %>% 
  rbind(
    data.frame(average_coverage_depth = c(9.9, 6.6, 11.5, 11.0, 7.7, 13.4, 4.3, 7.8, 8.8, 9.0, 13.4, 0.3, 17.6, 3.0),
           bases_with_coverage = c(66.6, 55.8, 62.7, 65.0, 56.6, 77.0, 70.5, 76.5, 60.0, 64.0, 73.8, 9.4, 62.8, 50.6)) %>% 
           add_column(data = "ZB") %>% 
           add_column(contig = 1:14) %>% 
           mutate(contig = as.factor(contig))) %>% 
           mutate_if(is.character, as.factor) %>% 
  ggplot(mapping = aes(x = contig, y = bases_with_coverage, colour = data)) +
  geom_col() + 
  facet_wrap(~data)
```

```{r}
library(gridExtra)
grid.arrange(coverage_plot1, coverage_plot2, bases_plot1, bases_plot2)
```


# Compare read depth between direct and indirect alignments for ZB data on the same sample (between contigs)

average_coverage_depth = c(9.9, 6.6, 11.5, 11.0, 7.7, 13.4, 4.3, 7.8, 8.8, 9.0, 13.4, 0.3, 17.6, 3.0),
bases_with_coverage = c(66.6, 55.8, 62.7, 65.0, 56.6, 77.0, 70.5, 76.5, 60.0, 64.0, 73.8, 9.4, 62.8, 50.6)

bases_with_coverage = c(38.4, 35.7, 43.5, 43.1, 36.5, 61.3, 33.2, 53.6, 34.1, 35.9, 55.5, 40.1, 26.0),
average_coverage_depth = c(3.2, 2.3, 3.8, 5.2, 3.1, 4.8, 0.9, 2.6, 1.8, 2.4, 4.8, 2.5, 0.1, 4.8, 0.7)

column for direct vs indirect 

```{r}
data.frame(average_coverage_depth = c(9.9, 6.6, 11.5, 11.0, 7.7, 13.4, 4.3, 7.8, 8.8, 9.0, 13.4, 0.3, 17.6, 3.0),
           bases_with_coverage = c(66.6, 55.8, 62.7, 65.0, 56.6, 77.0, 70.5, 76.5, 60.0, 64.0, 73.8, 9.4, 62.8, 50.6)) %>% 
           add_column(alignment = "Direct") %>%
           add_column(contig = 1:14) %>% 
  rbind(
data.frame(bases_with_coverage = c(38.4, 35.7, 43.5, 43.1, 36.5, 61.3, 33.2, 53.6, 34.1, 35.9, 55.5, 2.5, 40.1, 26.0),
           average_coverage_depth = c(3.2, 2.3, 3.8, 5.2, 3.1, 4.8, 0.9, 2.6, 1.8, 2.4, 4.8, 0.1, 4.8, 0.7)) %>% 
           add_column(alignment = "Indirect") %>%
           add_column(contig = 1:14)) %>% 
  ggplot(mapping = aes(x = alignment, y = average_coverage_depth, colour = alignment)) +
  geom_col() +
  facet_wrap(~contig) +
  theme(axis.title.x = element_blank()) +
  ggtitle("Average coverage: direct vs indirect alignment (ZB sample 09), faceted by contigs")
```


# Using samtools depth summary

Here I group by the contigs and get the average read depth at each base - this could be wrong. Tablet gives percentage of bases with coverage, and then average coverage depth. Thie coverage depth may only be for those bases with coverage > 0.

Thus try converting all zeroes to NA and then use the same summarise but with na.rm = T.


```{r}
read_depth_data <- function(file_path, dataset, alignment){
read_tsv(file_path, col_names = c("Contig", "Bases", "S01", "S02", "S03", "S04", "S05", "S06", "S07", "S08", "S09", "S10", "S11", "S12", "S13")) %>% 
  select(!Bases) %>% 
  filter(grepl("ordered", Contig)) %>% 
  group_by(Contig) %>% 
  summarise_all(mean) %>%
  t() %>% 
  as.data.frame() %>% 
  row_to_names(1) %>% 
  add_column(Data = dataset, Alignment = alignment) %>% 
  rownames_to_column("Sample")
}

depth_summary <- read_depth_data("data/read_depth/ZB_direct_depth_summary.tsv", "ZB", "Direct") %>% 
  rbind((read_depth_data("data/read_depth/ZB_indirect_depth_summary.tsv", "ZB", "Indirect"))) %>% 
  rbind((read_depth_data("data/read_depth/S100_direct_depth_summary.tsv", "Sanger", "Direct"))) %>%
  rbind((read_depth_data("data/read_depth/S100_indirect_depth_summary.tsv", "Sanger", "Indirect"))) %>% 
  rbind((read_depth_data("data/read_depth/SS_direct_depth_summary.tsv", "Initial Subset", "Direct"))) %>% 
  rbind((read_depth_data("data/read_depth/SS_indirect_depth_summary.tsv", "Initial Subset", "Indirect"))) %>%
  mutate_if(is.factor, as.character) %>% 
  mutate_at(grep("ordered", colnames(.)), as.numeric)
```

## Plot

```{r}
depth_summary %>% 
  ggplot() +
  geom_point(mapping = aes(x = Sample, y = ordered_PKNH_01_v2, colour = Data, shape = Alignment)) +
  scale_y_log10()
```

## With zeroes removed

```{r}
read_depth_data_na <- function(file_path, dataset, alignment){
read_tsv(file_path, col_names = c("Contig", "Bases", "S01", "S02", "S03", "S04", "S05", "S06", "S07", "S08", "S09", "S10", "S11", "S12", "S13")) %>% 
  select(!Bases) %>% 
  filter(grepl("ordered", Contig)) %>% 
  na_if(0) %>% 
  group_by(Contig) %>% 
  summarise_all(mean, na.rm=T) %>%
  t() %>% 
  as.data.frame() %>% 
  row_to_names(1) %>% 
  add_column(Data = dataset, Alignment = alignment) %>% 
  rownames_to_column("Sample")
}

depth_summary_no_na <- read_depth_data_na("data/read_depth/ZB_direct_depth_summary.tsv", "ZB", "Direct") %>% 
  rbind((read_depth_data_na("data/read_depth/ZB_indirect_depth_summary.tsv", "ZB", "Indirect"))) %>% 
  rbind((read_depth_data_na("data/read_depth/S100_direct_depth_summary.tsv", "Sanger", "Direct"))) %>%
  rbind((read_depth_data_na("data/read_depth/S100_indirect_depth_summary.tsv", "Sanger", "Indirect"))) %>% 
  rbind((read_depth_data_na("data/read_depth/SS_direct_depth_summary.tsv", "Initial Subset", "Direct"))) %>% 
  rbind((read_depth_data_na("data/read_depth/SS_indirect_depth_summary.tsv", "Initial Subset", "Indirect"))) %>%
  mutate_if(is.factor, as.character) %>% 
  mutate_at(grep("ordered", colnames(.)), as.numeric)

depth_summary_no_na %>% 
  ggplot() +
  geom_point(mapping = aes(x = Sample, y = ordered_PKNH_01_v2, colour = Data, shape = Alignment)) +
  scale_y_log10()
```

## Just samples SS13 and ZB31 (what we looked at in tablet)

From tablet:
SS average coverage = 2.6, 3.4, 28.3, 3.4, 3.3, 7.9, 274.2, 62.2, 3.1, 61.3, 9.0, 0.035, 5.5, 19.2
SS bases with coverage (%) = 54.1, 54.4, 70.1, 47.0, 56.2, 80.7, 99.8, 89.0, 46.6, 53.5, 75.3, 1.9, 57.3, 77.4

ZB average coverage = 124.0, 115.6, 123.6, 137.6, 105.9, 139.7, 42, 106.2, 129.1, 97.4, 140.4, 4.5, 176.8, 33.9
ZB bases with coverage (%) = 84.7, 75.5, 83.4, 80.9, 75.7, 91.4, 99.5, 94.3, 79.1, 83.1, 85.1, 14.6, 86.6, 79.6

```{r}
depth_summary %>%
  filter(Sample == "S11", Data == "ZB", Alignment =="Direct") %>% 
  rbind(depth_summary %>% 
    filter(Sample == "S12", Data == "Initial Subset", Alignment =="Direct")) %>% 
  select(2:16) %>% 
  t() %>% 
  as.data.frame() %>% 
  rownames_to_column("Contig") %>% 
  rename("ZB "= V1, "Subset" = V2) %>%  
  filter(grepl("ordered", Contig)) %>% 
  pivot_longer(cols = !Contig, names_to = "Data", values_to = "Depth") %>%  
  mutate(Depth = as.character(Depth)) %>% 
  mutate(Depth = as.numeric(Depth)) %>% View()
  ggplot() +
  geom_point(mapping = aes(x = Contig, y = Depth, colour = Data)) +
  theme(axis.text.x = element_blank())
```

### Without NA
```{r}
depth_summary_no_na %>%
  filter(Sample == "S11", Data == "ZB", Alignment =="Direct") %>% 
  rbind(depth_summary %>% 
    filter(Sample == "S12", Data == "Initial Subset", Alignment =="Direct")) %>% 
  select(2:16) %>% 
  t() %>% 
  as.data.frame() %>% 
  rownames_to_column("Contig") %>% 
  rename("ZB "= V1, "Subset" = V2) %>%  
  filter(grepl("ordered", Contig)) %>% 
  pivot_longer(cols = !Contig, names_to = "Data", values_to = "Depth") %>%  
  mutate(Depth = as.character(Depth)) %>% 
  mutate(Depth = as.numeric(Depth)) %>% 
  ggplot() +
  geom_point(mapping = aes(x = Contig, y = Depth, colour = Data)) +
  theme(axis.text.x = element_blank())
```

# Average read depth at each contig for each dataset 

```{r}
depth_summary %>% 
  select(!1) %>% 
  group_by(Data, Alignment) %>% 
  summarise_all(mean) %>% 
  pivot_longer(cols = !c(Data, Alignment), names_to = "Contig", values_to = "Depth") %>% 
  ggplot() +
  geom_point(mapping = aes(x = Contig, y = Depth, colour = Data, shape = Alignment)) +
  theme(axis.text.x = element_blank(), axis.ticks = element_blank()) +
  ggtitle("Average read depth at each contig for each dataset")
```

## With zeroes removed

```{r}
depth_summary_no_na %>% 
  select(!1) %>% 
  group_by(Data, Alignment) %>% 
  summarise_all(mean) %>% 
  pivot_longer(cols = !c(Data, Alignment), names_to = "Contig", values_to = "Depth") %>% 
  ggplot() +
  geom_point(mapping = aes(x = Contig, y = Depth, colour = Data, shape = Alignment)) +
  theme(axis.text.x = element_blank(), axis.ticks = element_blank()) +
  ggtitle("Average read depth at each contig (only aligned bases) for each dataset ")
```

## Calculate the percentage of NA values (bases with no coverage) at each contig and within each sample

```{r}
base_pairs <- function(file_path, dataset, alignment){
read_tsv(file_path, col_names = c("Contig", "Bases", "S01", "S02", "S03", "S04", "S05", "S06", "S07", "S08", "S09", "S10", "S11", "S12", "S13")) %>% 
  select(!Bases) %>% 
  filter(grepl("ordered", Contig)) %>% 
  na_if(0) %>% 
  group_by(Contig) %>% 
  summarise_all(funs(sum(is.na(.))/length(.) * 100)) %>% 
  t() %>% 
  as.data.frame() %>% 
  row_to_names(1) %>% 
  add_column(Data = dataset, Alignment = alignment) %>% 
  rownames_to_column("Sample")
}


base_pair_summary <- base_pairs("data/read_depth/ZB_direct_depth_summary.tsv", "ZB", "Direct") %>% 
  rbind((base_pairs("data/read_depth/ZB_indirect_depth_summary.tsv", "ZB", "Indirect"))) %>% 
  rbind((base_pairs("data/read_depth/S100_direct_depth_summary.tsv", "Sanger", "Direct"))) %>%
  rbind((base_pairs("data/read_depth/S100_indirect_depth_summary.tsv", "Sanger", "Indirect"))) %>% 
  rbind((base_pairs("data/read_depth/SS_direct_depth_summary.tsv", "Initial Subset", "Direct"))) %>% 
  rbind((base_pairs("data/read_depth/SS_indirect_depth_summary.tsv", "Initial Subset", "Indirect"))) %>%
  mutate_if(is.factor, as.character) %>% 
  mutate_at(grep("ordered", colnames(.)), as.numeric)
```

# Average NA at each contig for each dataset 

```{r}
base_pair_summary %>% 
  select(!1) %>% 
  group_by(Data, Alignment) %>% 
  summarise_all(mean) %>% 
  pivot_longer(cols = !c(Data, Alignment), names_to = "Contig", values_to = "Bases") %>% 
  ggplot() +
  geom_point(mapping = aes(x = Contig, y = Bases, colour = Data, shape = Alignment)) +
  theme(axis.text.x = element_blank(), axis.ticks = element_blank()) +
  ggtitle("Average percentage of bases WITHOUT coverage for each contig")
```

Despite obious trends showing ZB data is on top, and some pretty good looking averages, when looking at the raw data there are some SHIT samples. Hopefully looking at the parasite load will shed some light here.


# Read in metadata and combine with summary data

```{r}
read_metadata <- function(file_path){
  
# select desired samples fromt the metadata
temp1 <- readxl::read_excel(file_path) %>% 
  filter(sampleid == "PK_SB_DNA_006" |
         sampleid == "PK_SB_DNA_008" |
         sampleid == "PK_SB_DNA_009" |
         sampleid == "PK_SB_DNA_011" |
         sampleid == "PK_SB_DNA_012" |
         sampleid == "PK_SB_DNA_014" |
         sampleid == "PK_SB_DNA_015" |
         sampleid == "PK_SB_DNA_019" |
         sampleid == "PK_SB_DNA_023" |
         sampleid == "PK_SB_DNA_030" |
         sampleid == "PK_SB_DNA_031" | 
         sampleid == "PK_SB_DNA_033" |
         sampleid == "PK_SB_DNA_034") %>% 
  add_column(Sample = 1:13) 

# change levels of subset to match depth/base data
temp2 <- temp1 %>% 
  slice(1:9) %>% 
  mutate(Sample = as.character(Sample)) %>% 
  mutate(Sample = paste0("S0", Sample)) 

# change levels of subset to match depth/base data
temp3 <- temp1 %>% 
  slice(10:13) %>% 
  mutate(Sample = as.character(Sample)) %>% 
  mutate(Sample = paste0("S", Sample)) 

# combine 
temp4 <- temp2 %>% rbind(temp3)

# filter depth/base data for ZB data, join them together and then combine with the metadata wrangled above
depth_summary %>% 
  filter(Data == "ZB") %>% 
  pivot_longer(cols = !c(Sample, Data, Alignment), names_to = "Contig", values_to = "Depth") %>%  
  left_join(
    (base_pair_summary %>% 
       filter(Data == "ZB") %>% 
       pivot_longer(cols = !c(Sample, Data, Alignment), names_to = "Contig", values_to = "Bases"))) %>% 
  right_join(temp4)
  
}


ZB_metadata <- read_metadata("data/metadata/PK_Sabah_Sample_naming_indexes.xlsx")
```

## Plot depth of each sample (direct alignment) at each contig

```{r}
ZB_metadata %>% 
  filter(Alignment == "Direct") %>% 
  ggplot(mapping = aes(x = Sample, y = Depth)) +
  geom_col() +
  facet_wrap(~Contig) +
  ggtitle("Depth of each sample (direct alignment) at each contig") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
```

Why does sample 11 have much higher read depth? Parasitemia?


## Parasitemia per sample at each contig

```{r}
ZB_metadata %>% 
  filter(Alignment == "Direct") %>% 
  ggplot(mapping = aes(x = Sample, y = parasitemia)) +
  geom_col() +
  facet_wrap(~Contig) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  ggtitle("Parasitemia of each sample (direct alignment)") 
```



## Plot relationship between average depth (across contigs) and parasitemia
 
```{r}
ZB_metadata %>%
  group_by(Sample) %>% 
  summarise_if(is.numeric, mean) %>% 
  ggplot(mapping = aes(x = parasitemia, y = Depth)) +
  geom_point() +
  scale_y_log10() +
  geom_smooth(method = "lm", se = T)
```

## Plot relationship between depth across samples in relation to parasitemia, faceted by contig

```{r}
ZB_metadata %>%
  filter(Sample != "S11") %>% 
  ggplot(mapping = aes(x = parasitemia, y = Depth)) +
  geom_point() +
  scale_y_log10() +
  geom_smooth(method = "lm", se = T) +
  facet_wrap(~Contig) +
  ylab("Depth (log10)")
```

## Plot relationship between depth across samples in relation to NA bases, faceted by contig

```{r}
ZB_metadata %>%
  filter(Sample != "S11") %>% 
  ggplot(mapping = aes(x = parasitemia, y = Bases)) +
  geom_point() +
  scale_y_log10() +
  geom_smooth(method = "lm", se = T) +
  facet_wrap(~Contig) +
  ylab("Bases (log10)")
```

## Plot relationship between depth and disease severity, faceted by contig

```{r}
ZB_metadata %>%
  filter(Sample != "S11") %>% 
  ggplot(mapping = aes(x = severe, y = Depth)) +
  geom_point() +
  geom_boxplot() +
  scale_y_log10() +
  geom_smooth(method = "lm", se = T) +
  facet_wrap(~Contig) +
  ylab("Depth (log10)")
```


# Export data to be read into Rmd easily for reports
```{r}
write_csv(depth_summary, "data/depth_summary.csv")
write_csv(base_pair_summary, "data/base_pair_summary.csv")
write_csv(ZB_metadata, "data/ZB_metadata.csv")
```



# Comparing our samples to those in previous studies

```{r}
depth_summary_updated <- depth_summary %>% 
  rbind(
    read_tsv("data/read_depth/previous_Pk_data.tsv", col_names = c("Contig", "Bases", "S01", "S02", "S03", "S04")) %>% 
       select(!Bases) %>% 
       filter(grepl("ordered", Contig)) %>% 
      group_by(Contig) %>% 
      summarise_all(mean) %>%
      t() %>% 
      as.data.frame() %>% 
      row_to_names(1) %>% 
      add_column(Data = "Previous_Pk_data", Alignment = "Direct") %>% 
      rownames_to_column("Sample")) %>%
  mutate_if(is.factor, as.character) %>% 
  mutate_at(grep("ordered", colnames(.)), as.numeric) 

depth_summary_updated %>% 
  select(!1) %>% 
  group_by(Data, Alignment) %>% 
  summarise_all(mean) %>% 
  pivot_longer(cols = !c(Data, Alignment), names_to = "Contig", values_to = "Depth") %>% 
  ggplot() +
  geom_point(mapping = aes(x = Contig, y = Depth, colour = Data, shape = Alignment)) +
  theme(axis.text.x = element_blank(), axis.ticks = element_blank()) 
```

```{r}
depth_summary_updated %>% 
  filter(Data == "Previous_Pk_data") %>% 
  select(1:15) %>% 
  pivot_longer(cols = !Sample, names_to = "Contig", values_to = "Depth") %>%  
  ggplot(mapping = aes(x = Sample, y = Depth)) +
  geom_col() +
  facet_wrap(~Contig) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
```

**The third sample seems like a bit of an outlier, so we can try plotting the averages without.**

```{r}
depth_summary_updated %>% 
  slice(-81) %>% 
  select(!1) %>%
  group_by(Data, Alignment) %>% 
  summarise_all(mean) %>% 
  pivot_longer(cols = !c(Data, Alignment), names_to = "Contig", values_to = "Depth") %>% 
  ggplot() +
  geom_point(mapping = aes(x = Contig, y = Depth, colour = Data, shape = Alignment)) +
  theme(axis.text.x = element_blank(), axis.ticks = element_blank()) 
```


```{r}
base_pair_summary_updated <- base_pair_summary %>% 
  rbind(
    read_tsv("data/read_depth/previous_Pk_data.tsv", col_names = c("Contig", "Bases", "S01", "S02", "S03", "S04")) %>% 
       select(!Bases) %>% 
       filter(grepl("ordered", Contig)) %>% 
      na_if(0) %>%
      group_by(Contig) %>% 
  summarise_all(funs(sum(is.na(.))/length(.) * 100)) %>% 
      t() %>% 
      as.data.frame() %>% 
      row_to_names(1) %>% 
      add_column(Data = "Previous_Pk_data", Alignment = "Direct") %>% 
      rownames_to_column("Sample")) %>%
  mutate_if(is.factor, as.character) %>% 
  mutate_at(grep("ordered", colnames(.)), as.numeric) 

base_pair_summary_updated %>% 
  select(!1) %>% 
  group_by(Data, Alignment) %>% 
  summarise_all(mean) %>% 
  pivot_longer(cols = !c(Data, Alignment), names_to = "Contig", values_to = "Bases") %>% 
  ggplot() +
  geom_point(mapping = aes(x = Contig, y = Bases, colour = Data, shape = Alignment)) +
  theme(axis.text.x = element_blank(), axis.ticks = element_blank()) +
  ggtitle("Average percentage of bases WITHOUT coverage for each contig")
```

# Trim Test

This data is an updated dataset of the ZB data. There was an issue in the trimming step previously (which I did not realise until way later) and so I re-ran the trimming and mapping steps. 

```{r}
depth_summary_updated_ZB_redo <- depth_summary_updated %>% 
  rbind(
    read_tsv("data/read_depth/ZB_redo_direct_depth_summary.tsv", 
    col_names = c("Contig", "Bases", "S01", "S02", "S03", "S04", "S05", "S06", "S07", "S08", "S09", "S10", "S11")) %>% 
  select(!Bases) %>% 
  filter(grepl("ordered", Contig)) %>% 
  group_by(Contig) %>% 
  summarise_all(mean) %>%
  t() %>% 
  as.data.frame() %>% 
  row_to_names(1) %>% 
  add_column(Data = "ZB_Redo", Alignment = "Direct") %>% 
  rownames_to_column("Sample") %>%
  mutate_if(is.factor, as.character) %>% 
  mutate_at(grep("ordered", colnames(.)), as.numeric)) %>% 
  rbind(
    read_tsv("data/read_depth/Finish_ZB_redo_direct_depth_summary.tsv", 
    col_names = c("Contig", "Bases", "S12", "S13")) %>% 
  select(!Bases) %>% 
  filter(grepl("ordered", Contig)) %>% 
  group_by(Contig) %>% 
  summarise_all(mean) %>%
  t() %>% 
  as.data.frame() %>% 
  row_to_names(1) %>% 
  add_column(Data = "ZB_Redo", Alignment = "Direct") %>% 
  rownames_to_column("Sample") %>%
  mutate_if(is.factor, as.character) %>% 
  mutate_at(grep("ordered", colnames(.)), as.numeric))



depth_summary_updated_ZB_redo %>% 
  select(!1) %>% 
  group_by(Data, Alignment) %>% 
  summarise_all(mean) %>% 
  pivot_longer(cols = !c(Data, Alignment), names_to = "Contig", values_to = "Depth") %>% 
  ggplot() +
  geom_point(mapping = aes(x = Contig, y = Depth, colour = Data, shape = Alignment)) +
  theme(axis.text.x = element_blank(), axis.ticks = element_blank()) +
  ggtitle("Average read depth per contig")
```


## Bases

```{r}
base_pair_summary_updated_ZB_redo <- base_pair_summary_updated %>% 
  rbind(
    read_tsv("data/read_depth/ZB_redo_direct_depth_summary.tsv", 
    col_names = c("Contig", "Bases", "S01", "S02", "S03", "S04", "S05", "S06", "S07", "S08", "S09", "S10", "S11")) %>% 
       select(!Bases) %>% 
       filter(grepl("ordered", Contig)) %>% 
      na_if(0) %>%
      group_by(Contig) %>% 
  summarise_all(funs(sum(is.na(.))/length(.) * 100)) %>% 
      t() %>% 
      as.data.frame() %>% 
      row_to_names(1) %>% 
      add_column(Data = "ZB_Redo", Alignment = "Direct") %>% 
      rownames_to_column("Sample") %>%
  mutate_if(is.factor, as.character) %>% 
  mutate_at(grep("ordered", colnames(.)), as.numeric)) %>% 
  rbind(
  read_tsv("data/read_depth/Finish_ZB_redo_direct_depth_summary.tsv", 
    col_names = c("Contig", "Bases", "S12", "S13")) %>% 
       select(!Bases) %>% 
       filter(grepl("ordered", Contig)) %>% 
      na_if(0) %>%
      group_by(Contig) %>% 
  summarise_all(funs(sum(is.na(.))/length(.) * 100)) %>% 
      t() %>% 
      as.data.frame() %>% 
      row_to_names(1) %>% 
      add_column(Data = "ZB_Redo", Alignment = "Direct") %>% 
      rownames_to_column("Sample") %>%
  mutate_if(is.factor, as.character) %>% 
  mutate_at(grep("ordered", colnames(.)), as.numeric))

base_pair_summary_updated_ZB_redo %>% 
  select(!1) %>% 
  group_by(Data, Alignment) %>% 
  summarise_all(mean) %>% 
  pivot_longer(cols = !c(Data, Alignment), names_to = "Contig", values_to = "Bases") %>% 
  ggplot() +
  geom_point(mapping = aes(x = Contig, y = Bases, colour = Data, shape = Alignment)) +
  theme(axis.text.x = element_blank(), axis.ticks = element_blank()) +
  ggtitle("Average percentage of bases WITHOUT coverage for each contig")
```
